---
title: "map"
author: "Shuhan Song"
date: "12/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(beepr)
library(sf)
library(gganimate)
library(transformr)
library(here)
library(tidyverse)
library(janitor)
```

### 6. A little map teaser

our first map with `{sf}` package

`sf`: by Edzer pebesma
Great beacause: sticky geometries

Data source: 
[https://nid.sec.usace.army.mil/ords/f?p=105:19:14606625072661::NO:::](National Inventory of Dams)

Get and check out the CA dams data: 
```{r}
dams <- read_csv("ca_dams.csv") %>% 
  clean_names() %>% 
  drop_na(latitude) %>% # only keep dams that have latitue, year_completed and langitude information
  drop_na(longitude) %>% 
  drop_na(year_completed)
```

Then make sure R understands that latitude & longitude are spatial coordinates using `sf::st_as_sf()`:
```{r}
# converts our data frame to and sf pobject using st_as_Sf
dams_sf <- st_as_sf(dams, 
                    coords = c("longitude","latitude")
                    )

st_crs(dams_sf) <- 4326



# Check class:
class(dams_sf)
```

What does that mean? Now R recognizes this as spatial data. 

Cool! Now let's get an outline for California: 

# NOTE: GET SAME PROJECTIONS HERE! 
```{r}
ca_border <- read_sf(here::here("ca_state_border"), layer = "CA_State_TIGER2016")

```


### 7. A little gganimate teaser

Now show how dams have been added over time with `gganimate`!
```{r}
ggplot() +
  geom_sf(data = ca_border,
          fill = "pink",
          color = "white") +
  geom_sf(data = dams_sf, 
          size = 1.5,
          color = "gray50") +
  theme_void() +
  labs(title = 'Year: {round(frame_time,0)}') +
  transition_time(year_completed) +  # stay after show up
  shadow_mark()
# not show up in markdown
beep(4)
```

